---
title: DATA 606 Data Project Proposal
author: "Tage N Singh"
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data

```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

This project will examine the relationship between the confirmed COVID-19 cases and deaths for the United States between 01/22/2020 and 02/27/2021. The datasets provide county level data and popluations for each of the counties in the USA. 

### Cases 

**What are the cases, and how many are there?**

The cases are provided in a time series format in csv format


### Data collection 

**Describe the method of data collection.**



### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**




### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**



### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

